---
title: "Homework 6"
author: "Onyeka Isamah"
output: github_document
---

# Problem 1
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(p8105.datasets)
library(viridis)
library(patchwork)
library(modelr)
```

## Importing and cleaning the data 
- Converted numeric to factor where appropriate (`babysex`, `frace`, `mrace`, `malform`)
- Checked for missing data - doesn't not appear to be any missing data

```{r}
birthweight = 
  read_csv("./birthweight.csv")
```

```{r}
birthweight_clean = 
  birthweight %>%  janitor::clean_names() %>% 
  
mutate(
    babysex = factor(babysex),
    frace = factor(frace),
    mrace = factor(mrace),
    malform = factor(malform)
) 

#looking for missing data by summing any missing data points##
sum(is.na(birthweight_clean))
```

### Proposed regression model for birthweight
 *Modeling process*: Research suggests that the mother's height and weight impacts a baby's birth weight. Based on this hypothesized structure that possible varibles that underly birthweight,  I am proposing a regression model with the predictor variables `delwt`, mother's weight at delivery (pounds), `mheight`: mother’s height (inches), `wtgain`: mother’s weight gain during pregnancy (pounds). 

**Model:** 
```{r}
fit = lm(bwt ~ delwt + mheight + wtgain, data = birthweight_clean)
## using linear model ##
## obtaining p-value and tidying summary coefficient ##
fit %>% broom::tidy() %>% 
  knitr::kable(digits = 3)
```

## Model Diagnostics - plotting residuals 

- Residuals for the proposed model appear to be normally distributed, however the predictions appear to be skewed. 

```{r}
modelr:: add_residuals(birthweight_clean, fit)
modelr::add_predictions(birthweight_clean, fit)


plot1 =
  modelr:: add_residuals(birthweight_clean, fit) %>% 
  ggplot(aes(x = resid)) + geom_density() + xlim(-2000, 2000)


plot2 = modelr:: add_predictions(birthweight_clean, fit) %>% 
  ggplot(aes(x = pred)) + geom_density() + geom_density() + xlim(-4000, 4000)

plot1 + plot2
```

## Model comparison

```{r}
set.seed(10)

cv_df = crossv_mc(birthweight_clean, 100) %>% 
  mutate(tain = map(train,as_tibble),
         test = map(test, as_tibble)) %>%
  #creating other models, model 1 length at birth and gestational age as predictors (main effects only), model 2 head circumference, length, sex, and all interactions (including the three-way interaction)##
  
  mutate(
    my_model = map(train, ~lm(bwt ~ delwt + mheight + wtgain, data = birthweight_clean)),
    model_1 = map(train, ~lm(bwt ~ blength, gaweeks, data = .x)),
    model_2_int = map(train, ~lm(bwt~ bhead + blength + babysex + bhead*blength + bhead*babysex + blength*babysex +bhead*babysex, data = .x))
  )  %>% 
   #creating RMSE for the 3 models ##
  mutate(
    rmse_my_model = map2_dbl(my_model, test, ~rmse(model = .x, data = .y)),
    rmse_model_1 = map2_dbl(model_1, test, ~rmse(model = .x, data = .y)),
    rmse_model_2_int = map2_dbl(model_2_int, test,~ rmse(model = .x, data = .y)))
  
```

- **Plotting RMSE for the 3 models**:

 We can see that model 2, the model with the interaction terms, has the smallest RMSE values in comparison to my proposed model, and the model with just the main effects (model 1). Out of these 3 models, model 2 is the best model since it has the least prediction errors as indicated by the RMSE distribtuion plot. 
```{r}
cv_df %>% 
  select(starts_with('rmse')) %>% 
  pivot_longer(
    everything(),
    names_to = 'model',
    values_to = 'rmse',
    names_prefix = 'rmse_' ) %>% 
  mutate(model = fct_inorder(model)
  ) %>% 
## plot ##
ggplot(aes(x= model, y= rmse, color= model)) + geom_violin()
```


# Problem 2:

Importing data set: 
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```


- We’ll focus on a simple linear regression with `tmax` as the response and `tmin` as the predictor.
- We are interested in the distribution of two quantities estimated from these data: 
**r̂2and  log(β̂0∗β̂1)**

1.  Getting the estimates (r-squared and log for beta parameter estimates)
```{r}
bootstrap_df = weather_df %>%  
  bootstrap(n = 5000) %>% 
  mutate(
    models = map(strap, ~lm(tmax~tmin, data = .x)),
    results = map(models, broom::tidy),
    results_glance = map(models, broom::glance)) %>% 
  select(-strap, models) %>% 
unnest(results_glance) %>% ## getting r.squared values##
  select(.id, results, r.squared) %>% 
  unnest(results) %>% ##getting estimates##
  select(.id, r.squared, term, estimate) %>% 
  mutate(
term = recode(term, '(Intercept)' = "intercept_b0", tmin= "tmin_slope_b1")) %>% 
  pivot_wider(names_from = term, values_from = estimate) %>%  ##tidying output##
  mutate(log_betas = log(intercept_b0*tmin_slope_b1)) ##taking the log of the slope and intercept estimates)##
    bootstrap_df %>% head() %>%  knitr::kable()
```

2. Plotting the distribution of intercept estimates, slope estimates, r-squares values, and log beta estimates.
- *Description*: Using the bootstrap method for 5000 repeated samples, the distribution of all of the estimates appear to be normally distributed. 

```{r}
intercept_dist = ggplot(bootstrap_df, aes(x = intercept_b0)) + geom_histogram()

r_squared_dist =  ggplot(bootstrap_df, aes(x = r.squared)) + geom_histogram()
  
b1_dist =  ggplot(bootstrap_df, aes(x = tmin_slope_b1)) + geom_histogram()
  
log_beta_dist = ggplot(bootstrap_df, aes(x = log_betas)) + geom_histogram()

intercept_dist + r_squared_dist + b1_dist + log_beta_dist

```

3. Using the 5000 estimates, identify the .5% and 97.5% quantiles to provide a 95% confidence interval for r̂2
 and log(β̂0∗β̂1)


- 95% confidence interval for `r.squared` = .894, .927


- 95% confidence interval for `log_betas` = .1.966, 2.058
.
```{r}
bootstrap_df %>% 
  summarize(
            upper_r2 = quantile(r.squared, .975),
            lower_r2 = quantile(r.squared, .025),
            upper_log_betas = quantile(log_betas, .975),
            lower_log_betas = quantile(log_betas, .025)
              ) %>% knitr::kable(digits = 3)
```

